### 2013-09-19 intro.tex
typos julia

%\includegraphics[width=\linewidth]{python_soul}%
%\flushright%
%Randall Munroe, \emph{xkcd.com}
\end{minipage}%
\\[2cm]
In computer vision research, the goal is to automatically extract information from an image.
In particular discerning semantic information, that is interpreting an image, is a prominent
research topic.
While much process has been made in recent years, computer vision systems still lag behind
humans in most tasks that require semantic information. Often these tasks can be formulated
in terms of \emph{semantic classes}, meaning categories of parts, objects or scenes.
Examples include answering questions such as ``Is this a picture of a beach?'', ``How many
cars are in this image?'' or even ``What objects lie on the table?''.
These questions illustrate a range of possible tasks involving semantic categories,
from classifying images of single objects, over localization and counting of object classes,
to parsing a scene fully into objects and object classes, together with their relations.
%
While humans can distinguish tens of thousands of object classes, and have
little trouble in interpreting complex scenes, current methods are often
restricted to a much smaller number of classes, and only have limited
capabilities to model interactions or relations.  We believe that
\emph{context} is one of the most important cues when it comes to classifying
objects, and therefore understanding scenes. Consequently, we target dense
labeling of scenes, taking object relations into account.  The task of densely
labeling image pixels into classes is called object class segmentation or
semantic segmentation.
\pagebreak

@@

The main part of this thesis investigates the use of structured learning~\citep{taskar2003max, tsochantaridis2006large}
algorithms to the task of semantic image segmentation. Both topics have receive
much attention in the computer vision and machine learning communities
lately~\citep{ladicky2009associative, krahenbuhl2012efficient,
branson2013efficient, blake2011markov}.
Unfortunately, learning structured prediction in computer vision applications
is still little understood.
We focus on the use of conditional random fields~(CRFs), which have shown great
promise for computer vision applications. Using the paradigm of structural
support vector machines, it is possible to learn conditional random fields to
directly minimize a loss of interest. In particular, conditional random fields
allow to combine different cues, possibly produced by other approaches, in a
principled manner. One of the main difficulties with CRF approaches to computer
vision problems is, that context in images is usually represented as a
neighborhood graph.  These graphs, by nature, contain many cycles, making
inference intractable in general. Consequently, learning algorithms have to
rely on approximate inference, with often unknown consequences to learning.

There have been several previous studies on learning structural support vector
machines, and learning for conditional random fields. The impact of approximate
inference was first investigated by \citet{finley2008training}, applying
structural support vector machines to multi-label data. Later, different works
investigated how to combine approximate inference and learning in a single
framework. \citet{meshi2010learning, komodakis2011efficient}, and
\citet{hazan2010primal} approached the problem using duality, and formulate
learning and inference as a joint optimization problem.
\citet{stoyanov2011empirical}, and later \citet{jancsarylearning} and
\citet{krahenbuhlparameter} formulate learning structured prediction as
optimizing an \emph{prediction engine}, that takes into account all aspects of
the model, in particular the inference algorithm used.
In this work, on the other hand, we follow the well-established algorithms for
learning structural support vector machines, and investigate how we can use the
available inference algorithms to obtain good results within a reasonable
time-frame.

@@

\section{Thesis Outline}
Before we delve into semantic segmentation and object class recognition, we
first investigate a general clustering algorithm in Chapter~\ref{ch:itm}.
Clustering is an important step in most semantic segmentation pipelines, in at
least two places: bottom-up segmentation and creation of dictionaries for
feature learning. We introduce a novel, information theoretic algorithm that
compares favourably with algorithms from the literature.  While we do not apply
our algorithm to the task of bottom-up segmentation, this is a promising avenue
for future research.

We introducing an algorithm for  semi-supervised learning of object class
segmentation in Chapter~\ref{ch:semi_supervised}, motivated by the problem of
annotating training data for semantic segmentation.

>>>>

%\includegraphics[width=\linewidth]{python_soul}%
%\flushright%
%Randall Munroe, \emph{xkcd.com}
\end{minipage}%
\\[2cm]
In computer vision research, the goal is to automatically extract information from an image.
In particular discerning semantic information, that is interpreting an image, is a prominent
research topic.
While much progress has been made in recent years, computer vision systems still lag behind
humans in most tasks that require semantic information. Often these tasks can be formulated
in terms of \emph{semantic classes}, meaning categories of parts, objects or scenes.
Examples include answering questions such as ``Is this a picture of a beach?'', ``How many
cars are there in this image?'' or even ``What objects lie on the table?''.
These questions illustrate a range of possible tasks involving semantic categories,
from classifying images of single objects, over localization and counting of object classes,
to parsing a scene fully into objects and object classes, together with their relations.
%
While humans can distinguish tens of thousands of object classes, and have
little trouble in interpreting complex scenes, current methods are often
restricted to a much smaller number of classes, and only have limited
capabilities to model interactions or relations.  We believe that
\emph{context} is one of the most important cues when it comes to classifying
objects, and therefore understanding scenes. Consequently, we target dense
labeling of scenes, taking object relations into account.  The task of densely
labeling image pixels into classes is called object class segmentation or
semantic segmentation.
\pagebreak

@@

The main part of this thesis investigates the use of structured learning~\citep{taskar2003max, tsochantaridis2006large}
algorithms to the task of semantic image segmentation. Both topics have received
much attention in the computer vision and machine learning communities
lately~\citep{ladicky2009associative, krahenbuhl2012efficient,
branson2013efficient, blake2011markov}.
Unfortunately, learning structured prediction in computer vision applications
is still little understood.
We focus on the use of conditional random fields~(CRFs), which have shown great
promise for computer vision applications. Using the paradigm of structural
support vector machines, it is possible to learn conditional random fields to
directly minimize a loss of interest. In particular, conditional random fields
allow to combine different cues, possibly produced by other approaches, in a
principled manner. One of the main difficulties with CRF approaches to computer
vision problems is that context in images is usually represented as a
neighborhood graph.  These graphs, by nature, contain many cycles, making
inference intractable in general. Consequently, learning algorithms have to
rely on approximate inference, with often unknown consequences to learning.

There have been several previous studies on learning structural support vector
machines, and learning for conditional random fields. The impact of approximate
inference was first investigated by \citet{finley2008training}, applying
structural support vector machines to multi-label data. Later, different works
investigated how to combine approximate inference and learning in a single
framework. \citet{meshi2010learning, komodakis2011efficient}, and
\citet{hazan2010primal} approached the problem using duality, and formulate
learning and inference as a joint optimization problem.
\citet{stoyanov2011empirical}, and later \citet{jancsarylearning} and
\citet{krahenbuhlparameter} formulate learning structured prediction as
optimizing a \emph{prediction engine}, that takes into account all aspects of
the model, in particular the inference algorithm used.
In this work, on the other hand, we follow the well-established algorithms for
learning structural support vector machines, and investigate how we can use the
available inference algorithms to obtain good results within a reasonable
time-frame.

@@

\section{Thesis Outline}
Before we delve into semantic segmentation and object class recognition, we
first investigate a general clustering algorithm in Chapter~\ref{ch:itm}.
Clustering is an important step in most semantic segmentation pipelines, in at
least two places: bottom-up segmentation and creation of dictionaries for
feature learning. We introduce a novel information theoretic algorithm that
compares favourably with algorithms from the literature.  While we do not apply
our algorithm to the task of bottom-up segmentation, this is a promising avenue
for future research.

We introduce an algorithm for  semi-supervised learning of object class
segmentation in Chapter~\ref{ch:semi_supervised}, motivated by the problem of
annotating training data for semantic segmentation.

### 2013-09-19 main.tex
typos julia

\begin{center}
\Large\sffamily
\vspace*{1cm}
\textbf{{\huge Methods for Learning Structured Prediction in Semantic Segmentation of Natural Images}}\\[5mm]
Dissertation\\
zur\\
Erlangung des Doktorgrades (Dr.\ rer.\ nat.)\\
der\\
Mathematisch-Naturwissenschaftlichen Fakult\"at\\
der\\
Rheinischen Friedrich-Wilhelm Universit\"at Bonn\\
vorgelegt von\\
{\LARGE Andreas Christian M\"uller}\\
aus\\
Offenbach am Main\\
Bonn, September 2013
\end{center}
\end{titlepage}
%
%\thispagestyle{empty}
\clearpage
\pagenumbering{Roman}
\setcounter{page}{2}
\vspace*{5cm}
\Large\sffamily
\begin{center}
Angefertigt mit Genehmigung\\
der Mathematisch-Naturwissenschaftlichen Fakult\"at\\
der Rheinischen Friedrich-Wilhelms-Universit\"at Bonn\\
\end{center}
1. Gutachter Prof. Dr. Sven Behnke\\
2. Gutachter Prof. Dr. Reinhard Klein\\
Tag der Promotion:\\
Erscheinungsjahr: 2013\\
\end{spacing}
\tableofcontents
%\listoffigures
%\listoftables
\chapter*{Zusammenfassung}
\enlargethispage{10mm}
\selectlanguage{german}
Automatische Segmentierung und Erkennung von semantischen Klassen in
nat\"urlichen Bildern ist ein wichtiges offenes Problem des maschinellen Sehens.
In dieser Arbeit untersuchen wir drei m\"oglichen Ans\"atzen der Erkennung:
ohne \"Uberwachung, mit \"Uberwachung auf Ebene von Bildern und mit \"Uberwachung auf Ebene
von Pixeln. Diese Arbeit setzt sich aus drei Teilen zusammen.

Im ersten Teil der Arbeit schlagen wir einen Clustering-Algorithmus vor,
der eine neuartige, informationstheoretische Zielfunktion optimiert. Wir zeigen, dass
der vorgestellte Algorithmus \"ublichen Standardverfahren aus der Literatur gegen\"uber
klare Vorteile auf vielen verschiedenen Datens\"atzen hat. Clustering ist ein wichtiger
Baustein in vielen Applikationen des machinellen Sehens, insbesondere in der
automatischen Segmentierung.

Der zweite Teil dieser Arbeit stellt ein Verfahren zur automatischen
Segmentierung und Erkennung von Objektklassen in nat\"urlichen Bildern vor, der
mit Hilfe von Supervision in Form von Klassen-Vorkommen auf Bildern in der Lage
ist ein Segmentierungsmodell zu Lernen. Nach unserem Wissen ist dies der erste
Ansatz dieser Art, der sich speziell auf komplexe Objektklassen bezieht.

Der dritte und letzte Teil der Arbeit untersucht einen der am weitesten
verbreiteten Ans\"atze zu semantischen Segmentierung und
Objektklassensegmentierung, Conditional Random Fields, verbunden mit Verfahren
der strukturierten Vorhersage.
%Die Vorhersage f\"ur Modelle zur Bild-Markierung
%mittels Conditional Random Fields ist f\"ur gew\"ohnlich nur approiximativ
%moeglich, bedingt d\"urch das Vorhandensein von Kreisen in den zugrunde
%liegenden Nachbarschaftsgraphen.  Aus diesem Grund, und trotz der Beliebtheit
%von Conditional Random Field Verfahren, ist das strukturierte Lernen f\"ur
%solche Anwendungen schlecht verstanden.
%
Wir untersuchen verschiedene Lernalgorithmen des struktuierten Lernens, insbesondere
im Zusammenhang mit approximativer Inferenz. Wir zeigen, dass es m\"oglich ist trotz des Vorhandenseins von
Kreisen in den Nachbarschaftsgraphen exakte strukturierte Modelle zur Bildsegmentierung zu lernen.
Mit den vorgestellten Methoden bringen den Stand der Kunst
auf zwei komplexen Datens\"atzen zur semantischen Segmentierung voran, dem MSRC-21 Datensatz
von RGB-Bildern und dem NYU V2 Datensatz von RGB-D Bildern von Innenraum-Szenen.

Wir stellen au{\ss}erdem eine Software-Bibliothek vor, die es erlaubt einen
weitreichenden Vergleich der besten Lernverfahren f\"ur strukturiertes Lernen
durchzuf\"uhren.
Unsere Studie erlaubt uns eine Charakterisierung der betrachteten Algorithmen
in einer Reihe von Anwendungen, insbesondere der semantischen Segmentierung und
Objektklassensegmentierung.

\selectlanguage{english}
\chapter*{Abstract}
Automatic segmentation and recognition of semantic classes in natural images is
an important open problem in computer vision. In this work, we investigates
three different approaches to recognition: without supervision, with
supervision on level of images, and with supervision on the level of pixels.
The thesis comprises three parts.

The first part introduces a clustering algorithm that optimizes a novel
information-theoretic objective function. We show that the proposed algorithm
has clear advantages over standard algorithm from the literature on a wide
array of datasets. Clustering algorithms are an important building block for
higher-level computer vision applications, in particular for semantic
segmentation.

>>>>

\begin{center}
\Large\sffamily
\vspace*{1cm}
\textbf{{\huge Methods for Learning Structured Prediction in Semantic Segmentation of Natural Images}}\\[5mm]
Dissertation\\
zur\\
Erlangung des Doktorgrades (Dr.\ rer.\ nat.)\\
der\\
Mathematisch-Naturwissenschaftlichen Fakult\"at\\
der\\
Rheinischen Friedrich-Wilhelm Universit\"at Bonn\\
vorgelegt von\\
{\LARGE Andreas Christian M\"uller}\\
aus\\
Offenbach am Main\\
Bonn, September 2013
\end{center}
\end{titlepage}
%
%\thispagestyle{empty}
\clearpage
\pagenumbering{Roman}
\setcounter{page}{2}
\vspace*{5cm}
\Large\sffamily
\begin{center}
Angefertigt mit Genehmigung\\
der Mathematisch-Naturwissenschaftlichen Fakult\"at\\
der Rheinischen Friedrich-Wilhelms-Universit\"at Bonn\\
\end{center}
1. Gutachter Prof. Dr. Sven Behnke\\
2. Gutachter Prof. Dr. Reinhard Klein\\
Tag der Promotion:\\
Erscheinungsjahr: 2013\\
\end{spacing}
\tableofcontents
%\listoffigures
%\listoftables
\chapter*{Zusammenfassung}
\enlargethispage{10mm}
\selectlanguage{german}
Automatische Segmentierung und Erkennung von semantischen Klassen in
nat\"urlichen Bildern ist ein wichtiges offenes Problem des maschinellen Sehens.
In dieser Arbeit untersuchen wir drei m\"oglichen Ans\"atze der Erkennung:
ohne \"Uberwachung, mit \"Uberwachung auf Ebene von Bildern und mit \"Uberwachung auf Ebene
von Pixeln.

Diese Arbeit setzt sich aus drei Teilen zusammen.
Im ersten Teil der Arbeit schlagen wir einen Clustering-Algorithmus vor,
der eine neuartige, informationstheoretische Zielfunktion optimiert. Wir zeigen, dass
der vorgestellte Algorithmus \"ublichen Standardverfahren aus der Literatur gegen\"uber
klare Vorteile auf vielen verschiedenen Datens\"atzen hat. Clustering ist ein wichtiger
Baustein in vielen Applikationen des machinellen Sehens, insbesondere in der
automatischen Segmentierung.

Der zweite Teil dieser Arbeit stellt ein Verfahren zur automatischen
Segmentierung und Erkennung von Objektklassen in nat\"urlichen Bildern vor, das
mit Hilfe von Supervision in Form von Klassen-Vorkommen auf Bildern in der Lage
ist ein Segmentierungsmodell zu lernen. Nach unserem Wissen ist dies der erste
Ansatz dieser Art, der sich speziell auf komplexe Objektklassen bezieht.

Der dritte und letzte Teil der Arbeit untersucht einen der am weitesten
verbreiteten Ans\"atze zur semantischen Segmentierung und
Objektklassensegmentierung, Conditional Random Fields, verbunden mit Verfahren
der strukturierten Vorhersage.
%Die Vorhersage f\"ur Modelle zur Bild-Markierung
%mittels Conditional Random Fields ist f\"ur gew\"ohnlich nur approiximativ
%moeglich, bedingt d\"urch das Vorhandensein von Kreisen in den zugrunde
%liegenden Nachbarschaftsgraphen.  Aus diesem Grund, und trotz der Beliebtheit
%von Conditional Random Field Verfahren, ist das strukturierte Lernen f\"ur
%solche Anwendungen schlecht verstanden.
%
Wir untersuchen verschiedene Lernalgorithmen des strukturierten Lernens, insbesondere
im Zusammenhang mit approximativer Inferenz. Wir zeigen, dass es m\"oglich ist trotz des Vorhandenseins von
Kreisen in den Nachbarschaftsgraphen exakte strukturierte Modelle zur Bildsegmentierung zu lernen.
Mit den vorgestellten Methoden bringen den Stand der Kunst
auf zwei komplexen Datens\"atzen zur semantischen Segmentierung voran, dem MSRC-21 Datensatz
von RGB-Bildern und dem NYU V2 Datensatz von RGB-D Bildern von Innenraum-Szenen.

\enlargethispage{15mm}
Wir stellen au{\ss}erdem eine Software-Bibliothek vor, die es erlaubt einen
weitreichenden Vergleich der besten Lernverfahren f\"ur strukturiertes Lernen
durchzuf\"uhren.
Unsere Studie erlaubt uns eine Charakterisierung der betrachteten Algorithmen
in einer Reihe von Anwendungen, insbesondere der semantischen Segmentierung und
Objektklassensegmentierung.

\selectlanguage{english}
\chapter*{Abstract}
Automatic segmentation and recognition of semantic classes in natural images is
an important open problem in computer vision. In this work, we investigate
three different approaches to recognition: without supervision, with
supervision on level of images, and with supervision on the level of pixels.
The thesis comprises three parts.

The first part introduces a clustering algorithm that optimizes a novel
information-theoretic objective function. We show that the proposed algorithm
has clear advantages over standard algorithms from the literature on a wide
array of datasets. Clustering algorithms are an important building block for
higher-level computer vision applications, in particular for semantic
segmentation.
